A geo-referenced micro-data set of real estate listings for Spain's three largest cities

### 2023-05-16 editorial-materials/EPB-2022-0116.R1_response_to_reviewer.Rmd

---
title: "A geo-referenced micro-data set of real estate listings for Spain’s three largest cities - Response to referees"
output: pdf_document
bibliography: bibEPB.bib
---

<!--
```{r install-data-package, include = FALSE}
# Run only once if needed to install data package `sobiEquity`
if (!require("idealista18", character.only = TRUE)) {
      devtools::install_github("https://github.com/paezha/idealista18")
  }
```

```{r setup,  cache=FALSE, warning = FALSE, include=FALSE, message = FALSE}
library(idealista18)
library(tidyverse)

set.seed(1234)
```
-->

Referee: 1

> Comments to the Author

> Dear Authors, Thank you for the opportunity to review this paper. I would appreciate it if you would highlight the changed text in the next submission. Otherwise it is difficult to identify which parts have changed. The unknown quality of this one year randomised data is the main issue for potential users and I repeat my suggestion that this data should be validated for the benefit potential readers and users. Below is a list of comments and questions which  need to be addressed.

> 1.Please clearly state the nature of the data in the abstract. It is worthwhile to mention the relocated and randomised dataset in the abstract. This house price dataset only has 189,923 records, which could be not really be described as a big dataset as you describe it in the abstract. Except the authors could provide an evidence/reference on why this is big dataset.

**Response:**

We appreciate your comment. There are a few publicly available datasets for hedonic price analysis, but they tend to be smaller. For example, @bonifaci2015real includes $n=1,042$ observations for Padua, in Italy; @delgiudice2018housing share a dataset with $n=576$ observations relating to rental prices in Naples, Italy; @solano2019dataset present a dataset with $n=1,623$ daily rental prices in Seville, Spain; and the dataset of @fuerst2020real includes $n=4,201$ property prices geocoded to the level of nine regions in England and Wales. The most similar in terms of geographical disaggregation and sample size is the dataset of @song2021hedonic which includes transactions for four cities in South Korea, namely Busan ($n=61,152$), Daegu ($n=32,363$), Daejeon ($n=21,114$), and Gwangju ($n=25,984$). In response to your comment we drop the terminology of "big data" and describe our dataset as "large" instead.

>2. Please check the typographical errors before your next submission. For example, the “thirdr” in your table1.

**Response:**

Thank you for highlighting this. We have carefully proof-read the document for typos and other errors.

>3. Please use part of your answer to my previous fourth question in the manuscript. Your answer is quite useful for your future users, especially those researchers who are not familiar with house price data in Spain. I quote your previous answer below:

> “idealista covers fairly well all segments in the Spanish market, both individual and professional advertisers. This dataset comprehends information about listing prices, thus it contains the market situation from the offer. Nevertheless, actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores. However listing prices reflect quite well the (transaction) reality of real estate markets using the offer perspective, they keep strong correlation between idealista and transaction prices can be established see Banco de España. Even though a great extent of correlation does exist between official transactional and asking records, both databases can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (ie. ad contacts or ad views) and price gaps. The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals. In September 2021, according to data from Similarweb (a site specialized in sites’ traffic volume comparison), in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic. Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 monthly million visits (57% of the total traffic) versus its immediate competitor with 19.9 million visits (19.3 % of total traffic). In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealist”

**Response:**

Thank you for this suggestion. We have adapted this answer to expand the text in the section "Data Description".

>4.For data description, it is worth showing solid evidence of the quality of this dataset. For example, what is the data coverage of this data compared to all the residential properties in the city?  Listing the advertised volume is not very helpful for users who wish to understand the quality.

Thank you. We appreciate your comment. To provide better context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} now shows the number of listings with respect to the total residential stock in each city (in 2018). Globally, the number of listings to the total housing stock is around 7%. Additional information about the number of free-market home purchases can be obtained from the Spanish National Institute. Globally, the number of listingss 2,7 times the number of transactions.

To provide some context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} shows the number of listings with respect to the total residential stock in each city in 2018. As seen in the table, the number of listings ranges between 6.1% of the total number of properties (in Madrid) and 8.1% (in Valencia). Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] shows that the number of listings in the {idealista18} package correspond to 81.3\% of recorded real estate transactions in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia.

```{=tex}
\begin{table}[!ht]
    \centering
      \caption{Total properties and transactionst three Spanish cities. Year 2018} 
  \label{tab:transactions} 
    \begin{tabular}{lccccc}
        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
        \hline
        Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
        Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
        Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
        \hline
        Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
        \hline
        \multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
        \multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
        \multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
    \end{tabular}
\end{table}
```

>5. Given the asking price is a resized version of the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences your quantitative research results and findings. Please provide critical evidence to explain how this randomised price will not influence academic research. For example, how does the house price variation at local level change after the price is randomised. To what degree will this influence a typical hedonic house price research approach? I am concerned about whether this randomised data could bias conclusions.

**Response:**

We appreciate the opportunity to explain this issue, as it may be of interest to academic researchers. Adding a small amount of random noise ($\pm 2.5$) to the prices has the effect of increasing the variance of estimates, but does not affect the mean. This is a well-known result in algebra of random variables.

Prices ($P$) in the dataset are calculated as the sum of two random variables, the raw price ($RP$) and some random noise ($\eta$) drawn from a uniform distribution as follows:
$$

$$

The mean of the raw price is:
$$
\mathrm{E}[RP] = \frac{1}{N}\sum_{i=1}^N R_i
$$

The mean of the white noise is zero by design:
$$
\mathrm{E}[\eta] = 0
$$

We know that the expectation of the sum of two random variables is:
$$
\mathrm{E}[P] = \mathrm{E}[RP + \eta] = \mathrm{E}[RP] + \mathrm{E}[\eta]
$$

\noindent and since $\mathrm{E}[\eta] = 0$ it follows that the prices are unbiased.

The variance of the sum of two random variables is:
$$
\mathrm{V}[P] = \mathrm{V}[RP + \eta] = \mathrm{V}[RP] + 2\mathrm{Cov}[RP,\eta] + \mathrm{V}[\eta]
$$

The variance of $\eta$ is:
$$
\mathrm{V}[\eta]=\frac{1}{12}\big(0.025 - (-0.25)\big)^2
$$

>6. In figure 1, is it possible to standardise this data rather than showing only the number of dwelling records in your house price data? I guess the city centre always has high advertised volumes but it also has a high density of dwellings. It is worthwhile to show values standardised by the number of dwellings.

**Response:** After this comment we included a new figure to display the percentage of listings relative to the total number of dwellings by district.

```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
library(idealista18)
library(dplyr)
library(ggplot2)
library(sf)
library(gridExtra)

data("properties_by_district")

# Barcelona
Barcelona_prop <- properties_by_district |>
  filter(CITY == "Barcelona") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Barcelona_prop$proportion)
pq <- round(q * 100, 2)

Barcelona_prop <- Barcelona_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot4 <- ggplot(data = Barcelona_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

# Madrid
Madrid_prop <- properties_by_district |>
  filter(CITY == "Madrid") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Madrid_prop$proportion)
pq <- round(q * 100, 2)

Madrid_prop <- Madrid_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot5 <- ggplot(data = Madrid_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

# Valencia
Valencia_prop <- properties_by_district |>
  filter(CITY == "Valencia") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Valencia_prop$proportion)
pq <- round(q * 100, 2)

Valencia_prop <- Valencia_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot6 <- ggplot(data = Valencia_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

# Arrange plots
grid.arrange(plot4, plot5, plot6, 
             ncol = 1, 
             nrow = 3)
```

---

*Dear Authors,* *Thank you for the opportunity to review this paper. I would appreciate it if you would highlight the changed text in the next submission. Otherwise it is difficult to identify which parts have changed. The unknown quality of this one year randomised data is the main issue for potential users and I repeat my suggestion that this data should be validated for the benefit potential readers and users. Below is a list of comments and questions which need to be addressed.*

> **Answer:** Thnak you.
> In the new version of the manuscrip, the changes are writen in red.

## First Comment

>Please clearly state the nature of the data in the abstract. It is worthwhile to mention the relocated and randomised dataset in the abstract. This house price dataset only has 189,923 records, which could be not really be described as a big dataset as you describe it in the abstract. Except the authors could provide an evidence/reference on why this is big dataset.

## Second Comment

2.  *Please check the typographical errors before your next submission. For example, the "thirdr" in your table1.*

> **Answer:** Thank you.
> A native English speaker has reviewed the document.

## Third Comment

3.  *Please use part of your answer to my previous fourth question in the manuscript. Your answer is quite useful for your future users, especially those researchers who are not familiar with house price data in Spain. I quote your previous answer below:*

*"dealista covers fairly well all segments in the Spanish market, both individual and professional advertisers. This dataset comprehends information about listing prices, thus it contains the market situation from the offer. Nevertheless, actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores. However listing prices reflect quite well the (transaction) reality of real estate markets using the offer perspective, they keep strong correlation between idealista and transaction prices can be established see Banco de España. Even though a great extent of correlation does exist between official transactional and asking records, both databases can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (ie. ad contacts or ad views) and price gaps. The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals. In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic. Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 monthly million visits (57% of the total traffic) versus its immediate competitor with 19.9 million visits (19.3 % of total traffic). In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealist"*

> **Answer:** Thank you. We include it.

## Fourth Comment

4.*For data description, it is worth showing solid evidence of the quality of this dataset. For example, what is the data coverage of this data compared to all the residential properties in the city? Listing the advertised volume is not very helpful for users who wish to understand the quality.*

> **Answer:** 
## Fifth Comment

5.  *Given the asking price is a resized version of the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences your quantitative research results and findings.* **Please provide critical evidence to explain how this randomised price will not influence academic research.** *For example, how does the house price variation at local level change after the price is randomised. To what degree will this influence a typical hedonic house price research approach? I am concerned about whether this randomised data could bias conclusions.*

> **Answer:** Of course, the results of any econometric model will be differents if the original data are used.
> Unfortunally, the data protection laws in Spain, specifically the Organic Law on the Protection of Personal Data and the Guarantee of Digital Rights (LOPDGDD) and the General Data Protection Regulation (GDPR) of the European Union, establish that personal data, including information on the location of a home, is protected and must be treated confidentially.
> It is not permitted to disclose the specific location of a home without the explicit consent of the owner or person who has the right to dispose of it.

> The small changes in the prices will be, in mean, a small impact on the results.

## Sixth Comment

6.  *In figure 1, is it possible to standardise this data rather than showing only the number of dwelling records in your house price data? I guess the city centre always has high advertised volumes but it also has a high density of dwellings. It is worthwhile to show values standardised by the number of dwellings*

```{r, eval = FALSE, include = FALSE, message=FALSE, fig.width=5, }
library(readxl)
library(idealista18)
library(dplyr)
library(ggplot2)
library(sf)
library("gridExtra")
Data <- read_excel("Adds_by_Distrits_idealista18.xlsx")
Data$Por <- Data$N*100/Data$N_CADASTRE
Barcelona_Polygons$District <- as.numeric(substr(Barcelona_Polygons$LOCATIONID,23,24))
Barcelona_District <- Barcelona_Polygons %>% group_by(District) %>% summarise(District = unique(District))
Madrid_Polygons$District <- as.numeric(substr(Madrid_Polygons$LOCATIONID,23,24))
Madrid_District <- Madrid_Polygons %>% group_by(District) %>% summarise(District = unique(District))
Valencia_Polygons$District <- as.numeric(substr(Valencia_Polygons$LOCATIONID,23,24))
Valencia_District <- Valencia_Polygons %>% group_by(District) %>% summarise(District = unique(District))

Data.bar <- Data[Data$CITY=="Barcelona",]
Data.mad <- Data[Data$CITY=="Madrid",]
Data.val <- Data[Data$CITY=="Valencia",]

### 2023-05-16 EPB/main_EPB.Rmd
Edits in preparation for resubmission and initializing renv

---
title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
runninghead: author \emph{et al}.
author:
- name: 
abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on the idealista.com real estate website. The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest. Along with the real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"

keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
classoption:
  - Royal
  - times
bibliography: bibEPB.bib
bibliographystyle: sageh

@@

@@

Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature. Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market. But to really understand the characteristics of the housing market, it is essential to have information/data at the point level. Therefore, it is becoming common for spatial analysis of urban environments to be developed with geo- referenced micro-data sets [@lopez2015]. However, the availability of this type of open data at the point level is limited, and not many data sets contain latitude/longitude coordinates for each dwelling. In some cases, researchers have had to resort to web scraping processes to obtain the large volumes of information that permit robust analyses [@gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].  These  web  scraping  processes  can  include  missing  data, download errors, duplicate records, etc. Furthermore, the authors of this research do not generally share the data sets.

We are also witnessing a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021] using reproducible or replicable  research  [@paez2021open; @brunsdon2021opening]. But  to  work  openly in  science, it  is  necessary to  have  free software and open data. While great efforts have been made to make free software available to researchers (e.g., R or Python), data regarding real estate markets tend to be treated as confidential, and there are still few open micro-data sets of housing markets available [@Song2021].

To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open micro-data set of geo-referenced dwelling listings. The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain’s three largest cities. To date, this data product is the biggest open geo-referenced micro-data set of the housing market in Spain. Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors. However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.

This micro-data set is expected to be used as a benchmark dataset to test empirical methods in a reproducible fashion. Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others. The data can also be used to study the segmentation of housing submarkets and related topics such as the impact of suburban areas on house prices. The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest. In any case, the data might be easily extended by spatially joining other datasets that contain information at administrative, census tract, or postal code levels.

The data set is distributed as an R package, name {idealista18}, which can be accessed from a public Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}. The database idealista18 is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.

@@

The open data product {idealista18} is an R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia. For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the R package. The following subsections describe each object. A full description of the data is available in the help section of the package. We have strived to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the R ecosystem mean that documentation is easily searchable; distribution as an R package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.

@@

The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad. The dwelling listing has been included in the ‘idealista18’ package as an sf object [@Pebesma]. The name of the sf object containing the dwelling listing includes the name of the city, followed by '_Sale' (e.g., Madrid_Sale) and includes a total of 42 variables. Each sf object includes the complete set of listings corresponding to the four quarters of the year 2018. Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter. The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter. The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.

```{r number-ads, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
H <- cbind(H,rowSums(H))
text_tbl <- data.frame(
City = c("Barcelona","Madrid","Valencia"),
H,
check.names = FALSE)
colnames(text_tbl) <- c("City\\Quarter", "First", "Second ", "Thirdr", "Fourth","Total ads")
xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "4em", italic = FALSE) %>%
column_spec(2, width = "3em") %>%
column_spec(3, width = "3em", latex_column_spec = "c") %>%
column_spec(4, width = "3em", latex_column_spec = "c") %>%
column_spec(5, width = "3em", latex_column_spec = "c") %>%
column_spec(6, width = "2em", latex_column_spec = "c")
```

>>>>

---
title: A geo-referenced micro-data set of real estate listings for Spain's three largest cities
runninghead: author \emph{et al}.
author:
- name: 
abstract: "This article presents an open data product with large geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on the idealista.com real estate website. The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest. Along with the real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"

keywords: Housing prices; hedonic price analysis; idealista.com; geo-referenced data; point-level data; open data; Spain
classoption:
  - Royal
  - times
bibliography: bibEPB.bib
bibliographystyle: sageh

@@

```{r load-packages, include=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
library(gridExtra)
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
```

```{r load-data, include=FALSE}
# Listings
data("Barcelona_Sale")
data("Madrid_Sale")
data("Valencia_Sale")

# Polygons
data("Barcelona_Polygons")
data("Madrid_Polygons")
data("Valencia_Polygons")

# Points of interest
data("Barcelona_POIS")
data("Madrid_POIS")
data("Valencia_POIS")

# Districts
data("properties_by_district")
```

@@

Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature. Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market: to this end, it is essential to have information/data at the point level. Therefore, it is becoming common for spatial analysis of urban environments to be use micro-data, geo-referenced as points [@lopez2015]. In some cases, researchers have had to resort to webscraping processes to obtain data for research [e.g., @gupta2022take; @arbia2020spatial; @Li2019; @lopez2015]. Alas, webscraping is a chancy process prone to download errors, missing  data, duplicated records, etc. Furthermore, researchers do not always share their webscrapped datasets, which limits reproducibility of their research. As we witness a growing interest in openness and reproducibility in geographical data science [@arribasl2021editorial; @paez2021open; @brunsdon2021opening], it  becomes increasingly urgent to have open data products to support research [@arribas2021].

Some researchers have already responded to this need for publicly available, geo-referenced datasets to support open, reproducible research. A few such datasets are now available to support the analysis of real estate markets, but they are sometimes geo-referenced at the level of large geographical zones, such as @fuerst2020real, an open dataset that includes $n=4,201$ property prices geocoded to the level of nine regions in England and Wales. Other datasets are geo-coded as points, including @bonifaci2015real, which includes $n=1,042$ observations for Padua, in Italy; @delgiudice2018housing, who share a dataset with $n=576$ observations relating to rental prices in Naples, Italy; and @solano2019dataset present a dataset with $n=1,623$ daily rental prices in Seville, Spain.

To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open micro-data set of geo-referenced dwelling listings. The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain's three largest cities. To date, this data product is one of the largest open geo-referenced micro-data set of the housing market in the world. The most similar in terms of geographical disaggregation and sample size is the dataset of @song2021hedonic which includes transactions for four cities in South Korea, namely Busan ($n=61,152$), Daegu ($n=32,363$), Daejeon ($n=21,114$), and Gwangju ($n=25,984$). It is certainly the largest in Spain.

The data set has been supplied directly by Idealista, and therefore is clean and free of download errors. However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.

This micro-data set can be used to benchmark new analytical methods in a reproducible fashion. Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others. The data can also be used to study the segmentation of housing submarkets and related topics such as the impact of suburban areas on house prices. The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest. In any case, the data might be easily extended by spatially joining other datasets that contain information at administrative, census tract, or postal code levels.

The data set is distributed as an `R` package, named {idealista18}, which can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}. The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.

@@

The open data product {idealista18} is an `R` package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia. For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the package.

The data provider (idealista.com) is a leading real estate portal in Spain, on par with Fotocasa. Smaller portals include Habitaclia and Milanuncios, the latter focusing almost exclusively on individual (i.e., non-professional) advertisers. In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), there were a total of 103 million page views on real estate portals in Spain. Of this, 94% of traffic was concentrated on four portals, of which idealista was the most important, followed by Fotocasa, Habitaclia, and Pisos.com. Traffic is also highly concentrated, and idealista.com was the the leader by far with 58.6 monthly million visits (57% of the total traffic) compared to its closest competitor with 19.9 million visits (19.3 % of total traffic).

As a result of its share of advertisments, idealista.com covers fairly well all segments of the Spanish market, including both individual and professional advertisers. This dataset includes information about listing prices, and therefore represents the market situation from the perspective of asking prices. This is a necessary compromise in the present case, since actual sale prices are not publicly available, and the information can only be accessed by paying high fees to Colegio de Registradores. However listing prices reflect quite well the (transaction) reality of real estate markets, and correlations between idealista listings and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031). Official transactions and asking prices can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (i.e., ad contacts or ad views) and price gaps.

To provide some context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} shows the number of listings with respect to the total residential stock in each city in 2018. As seen in the table, the number of listings ranges between 6.1% of the total number of properties (in Madrid) and 8.1% (in Valencia). Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] shows that the number of listings in the {idealista18} package correspond to 81.3\% of recorded real estate transactions in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia.

```{=tex}
\begin{table}[!ht]
    \centering
      \caption{Total properties and transactionst three Spanish cities. Year 2018} 
  \label{tab:transactions} 
    \begin{tabular}{lccccc}
        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
        \hline
        Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
        Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
        Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
        \hline
        Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
        \hline
        \multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
        \multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
        \multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
    \end{tabular}
\end{table}
```

The following subsections describe each object. A full description of the data is available in the help section of the package. We have tried to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the R ecosystem mean that documentation is easily searchable; distribution as an R package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.

@@

The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad. The dwelling listing has been included in the {idealista18} package as a simple features (sf) object [@Pebesma] with point geometry in latitude and longitude. The name of each sf object with the list of dwellings is the name of the city followed by '_Sale' (e.g., Madrid_Sale). Each data object includes a total of 42 variables and the complete set of listings corresponding to the four quarters of 2018. Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter. The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter. The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.

```{r number-ads, echo = FALSE, message = FALSE}
H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
H <- cbind(H,rowSums(H))
text_tbl <- data.frame(
City = c("Barcelona","Madrid","Valencia"),
H,
check.names = FALSE)
colnames(text_tbl) <- c("City\\Quarter", "First", "Second ", "Third", "Fourth","Total ads")
xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") |>
xtable2kable(include.rownames = FALSE) |>
kable_styling(full_width = F) |>
column_spec(1, width = "4em", italic = FALSE) |>
column_spec(2, width = "3em") |>
column_spec(3, width = "3em", latex_column_spec = "c") |>
column_spec(4, width = "3em", latex_column_spec = "c") |>
column_spec(5, width = "3em", latex_column_spec = "c") |>
column_spec(6, width = "2em", latex_column_spec = "c")
```

Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the location of the dwelling. The exact location has been masked, as described in Section [Anonymizing the data set](#anonymizing)). Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable. The dwelling listings were enriched with a number of additional attributes from the Spanish cadastre [@Catastro]. The cadastral information is described in Table \ref{tab:variables}, with the prefix CAD in the variable name. The cadastral features were assigned by applying the features of the nearest parcel to the coordinates. The year the dwelling was built (CONSTRUCTIONYEAR) given by the advertiser was revised since the year of construction is entered on the website by users, and it is therefore subject to errors and incomplete data (40\% missing data). To resolve this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included, assigning the cadastral construction year from the nearest cadastral parcel whenever the value was outstanding (date was after publication date or year of construction was before 1500) or when the value supplied by the advertiser was missing.

Additionally, the distance of each dwelling to several points of interest was included in the sf object: distance to the city center, distance to the closest metro station, and distance to a major street (La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez for Valencia). The last rows of Table \ref{tab:variables} show the mean values of these variables.
